---
title: "Survey Analysis"
author: Pete Mohanty, PhD
date: "`r format(Sys.time(), '%B %d, %Y')`"
output:
  pdf_document: default
  word_document: default
  html_notebook: default
urlcolor: blue
params:
  spssfile: !r 1
  surveywave: !r 2016
---

```{r, echo = FALSE, message = FALSE, warning = FALSE}
library(pacman)                            
p_load(knitr, foreign, questionr, tidyverse)
opts_chunk$set(echo = FALSE)                                        # don't display R code,
opts_chunk$set(warning = FALSE)                                     # warnings,
opts_chunk$set(message = FALSE)                                     # messages,
opts_chunk$set(comment = "")                                        # or ## before each line of R output.
opts_chunk$set(results = "asis")                                    # enable kable() rendering inside loops, functions
knitr::opts_knit$set(root.dir = '/users/mohanty/Dropbox/git/ds101/automating/')
```


```{r}
datafiles <- dir(pattern = "sav", recursive = TRUE)
survey <- read.spss(datafiles[params$spssfile], to.data.frame = TRUE)
```

# `r params$surveywave` Survey

This is basic report of summary statistics.

Here's a weighted crosstab.

```{r}
kable(wtd.table(survey$ideo, survey$sex, survey$weight)/nrow(survey), digits = 2)
```

```{r}
# here is some inevitable data cleaning (shortening labels, etc.)

x <- names(survey)[grep("q2[[:digit:]]",  names(survey))]  
# grab all Qs 20, 21, 22... 29 including 21a, 21b, etc.   
y <- c("ideo", "party")

levels(survey[["ideo"]])[4] <- "Liberal"
levels(survey[["party"]])[4] <- "None"
levels(survey[["party"]])[5] <- "Other"

for(i in c(x, y)){
  
  levels(survey[[i]]) <- gsub("[^[:alnum:] ]", "", levels(survey[[i]])) 
  # gsub is find and replace (substitute)
  # above removes non-alphanumeric characters except space
  
  v <- grep("VOL", levels(survey[[i]]))   # keep track of voluntary responses...
  levels(survey[[i]]) <- gsub("VOL ", "", levels(survey[[i]]))
  levels(survey[[i]]) <- gsub("OR", "", levels(survey[[i]]))
  levels(survey[[i]])[grep("Refused", levels(survey[[i]]))] <- "DK"
  levels(survey[[i]])[grep("early", levels(survey[[i]]))] <- "Too early"
  levels(survey[[i]])[grep("Neither", levels(survey[[i]]))] <- "Other"
  
  for(j in v){
     levels(survey[[i]])[j] <- paste0(levels(survey[[i]])[j], "*") 
     # finishes replacing (VOL) with more compact *
  }
}

# collapse various types of spontaneous non-response (no party, don't know, etc.)
levels(survey$q1)[3] <- "Don't Know (VOL)"
tmp <- as.character(survey$party)
survey$party.clean <- ifelse(tmp %in% levels(survey$party)[1:3], tmp, "VOL: Other")

survey$race <- survey$racethn
levels(survey$race) <- gsub(" non-Hispanic", " (nH)", levels(survey$race)) 
levels(survey$race)[5] <- "DK*"

```

Presidential approval numbers that only look at approval among Democrats or Republicans since, in bad times for the President, people may stop identifying with either. Here is a graph that hopefully shows a more complete picture.

```{r}
PA <- ggplot(survey) + theme_minimal()
PA <- PA + geom_bar(aes(q1, y = (..count..)/sum(..count..), weight = weight, fill = q1))
# y = (..count..)/sum(..count..) makes it percentage rather than a count
PA <- PA + facet_grid(party.clean ~ .) + theme(strip.text.y = element_text(angle = 45))
# facet_grid is what makes a graph for each group
PA <- PA + xlab("") + ylab("Percent of Country") 
PA <- PA + ggtitle(paste("Presidential Approval", params$surveywave))
PA <- PA + scale_y_continuous(labels = scales::percent)
PA
```


# Multiple Crosstabs as Columns 

Here is the data displayed in different fashion (crosstabs as additional columns).

```{r}
source("https://raw.githubusercontent.com/rdrr1990/datascience101/master/automating/tabs.R")
kable(tabs(survey, "q1", c("sex", "race"), weight = "cellweight"))
```

### Additional Comparisons
Here's a whole bunch more. In each case q21, q22, ... q29 will be weighted and broken down by ideology and then party.


```{r}
for(i in x){
  for(j in y){
    cat("\nWeighted proportions for", i, "broken down by", j, "\n")
    print(kable(wtd.table(survey[[i]], survey[[j]], survey$weight)/nrow(survey), digits = 2))
    cat("\n")          # break out of table formatting
  }
  cat("\\newpage")
}

```
